---
title: "Factors influencing life expectancy"
author: "Group 11"
date: "4/18/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("tinytex")
#library(tinytex)
```

#installing the required packages
```{r}
#install.packages("rlang")
library(rlang)
#install.packages("kableExtra")
#install.packages("gridExtra")
```
```{r}
#install.packages("scales")
#install.packages("leaps")
#install.packages("GGally")
```
```{r}
#install.packages("MASS")
#install.packages("lmtest")
#install.packages("MLmetrics")
#install.packages("car")
```

#calling the libraries
```{r}
#install.packages("tidyverse")

#library(tidyverse) 
```
```{r}
#library(data.table)
#library(DT)
#library(kableExtra)
```
```{r}
library(gridExtra)
library(knitr)
library(tidyverse)
library(caret)
```

```{r}
#remove.packages("rlang")
```
```{r}
library(psych)
library(stats)
library(leaps)
#install.packages("ggplot2")
library(ggplot2)
library(GGally)
```
```{r}
#install.packages("rlang")

#library(rlang)
```
```{r}
library(MASS)
library(lmtest)
library(car)
library(MLmetrics)
```

#loading the dataset
```{r}
WHO <- read.csv("C:/Users/13176/Desktop/Life Expectancy Data.csv")
```

#data summary
```{r}
summary(WHO)
```
#descriptive statistics of the data
```{r}
library(psych)
describe(WHO)
head(WHO)
```
#converted the catergorical variable status to mumeric 
#assigned 1 for developed and 0 for devoloping countries
```{r}
sapply(WHO, class)
WHO$Life.expectancy<-as.numeric(WHO$Life.expectancy)

WHO$Status[WHO$Status=="Developed"]=1
WHO$Status[WHO$Status=="Developing"]=0
```
#box plots 
```{r}
library(DataExplorer)
plot_boxplot(WHO, "Status")
```
```{r}
library("ggplot2")
library(ggthemes)
WHO %>% 
         keep(is.numeric) %>% 
         gather() %>%           
         ggplot(aes(value)) +  
         facet_wrap(~key, scale = "free") + 
           geom_histogram(aes(y = ..density..), color = "darkblue", fill = "lightblue") + geom_density(alpha = 0.2) + theme_economist()
```
#LIFE EXPECTANCY VS INCOME COMPOSITIONS OF RESOURCES GRAPH
```{r}
plot(y = WHO$Life.expectancy,
     x = WHO$Income.composition.of.resources,
     main = "LIFE EXPECTANCY VS INCOME COMPOSITIONS OF RESOURCES ",
     xlab = "INCOME COMPOSITION",
     ylab = "LIFE EXPECTANCY",
     pch = 20,
     col = "RED")
abline(60,1,
       col = "maroon")
```
#LIFE EXPECTANCY VS SCHOOLING GRAPH
```{r}
plot(y = WHO$Life.expectancy,
     x = WHO$Schooling,
     main = "LIFE EXPECTANCY VS SCHOOLING",
     xlab = "SCHOOLING",
     ylab = "LIFE EXPECTANCY",
     pch = 20,
     col = "GREEN")
abline(50,1,
       col = "maroon")
```
#LIFE EXPECTANCY VS ADULT MORTALITY GRAPH
```{r}
plot(y = WHO$Life.expectancy,
     x = WHO$Adult.Mortality,
     main = "LIFE EXPECTANCY VS ADULT MORTALITY",
     xlab = "ADULT MORTALITY",
     ylab = "LIFE EXPECTANCY",
     pch = 20,
     col = "VIOLET")
abline(80,- 1,
       col = "maroon")
```
#LIFE EXPECTANCY VS POPULATION GRAPH
```{r}
plot(y = WHO$Life.expectancy,
     x = WHO$Population,
     main = "LIFE EXPECTANCY VS POPULATION",
     xlab = "POPULATION",
     ylab = "LIFE EXPECTANCY",
     pch = 20,
     col = "PURPLE")
abline(50,1,
       col = "maroon")
```

#  filtering the data and group the data by country to obtain the mean life expectancy of each country over the 16 years.

```{r}
developing<-WHO %>% filter(Status=="0")%>%group_by(Country) %>%summarise(Average_life=mean(Life.expectancy))

developed<-WHO%>% filter(Status=="1")%>%group_by(Country) %>%summarise(Average_life=mean(Life.expectancy))
```
# F-Test to check variance of the two populations
```{r}
var.test(developed$Average_life,developing$Average_life)

#The observed p-value is less than alpha(0.05) .Hence we reject the null hypothesis and accept the alternate statement that the variance of two populations is not equal.
```
#two-sample T-Test for the equality of the two means
```{r}
t.test(developed$Average_life,developing$Average_life, alternative= "greater",var.equal = FALSE)
#The null hypothesis is rejected against the alternative hypothesis as the p-value<0.05. Hence we conclude that life expectancy in developed countries is more than that of developing countries with 95% confidence.
```
#Interpretation of multiple linear regression analysis followed by regression diagnostics to identify issues with the model.
```{r}
mod2 = lm(Life.expectancy ~ Income.composition.of.resources + Schooling + Adult.Mortality + Population, data = WHO)
summary(mod2)
```
#Anova test
```{r}
anova(mod2)
```
#plots
```{r}
plot(mod2)
```
